Reasoning about a Capability Machine with Local Capabilities\\ Provably Safe Stack and Return Pointer Management (without OS Support)

### 2017-02-14 csf-2017/paper.tex
Added another version of figure 3 (in comment). Saves a bit of space, but is not as clear and lacks the 'failed' case of table
typeset programs so they have same font size as math
updated in text code to match the new style

\usepackage{cite}
\usepackage{amsmath,amsfonts,amssymb,amsthm}
\usepackage{thmtools}
\usepackage{stmaryrd}
\usepackage{natbib}
\usepackage{url}
\usepackage{array}
\usepackage{arydshln}
\usepackage{ifthen}
\usepackage{ifpdf}
\usepackage{verbatim}
\usepackage{mathpartir}
\usepackage{listings}
\usepackage{hyperref}
\usepackage{multicol}
\usepackage{natbib}
\lstset{
  basicstyle=\ttfamily,
  columns=fullflexible,
  keepspaces=true,
  mathescape
}

@@

\figurename~\ref{fig:op-sem} shows an excerpt of the operational semantics containing a number of representative
instructions. Essentially, a configuration $\Phi$ will either decode and execute
the instruction at $\memreg{\pcreg}$ if it is executable and the current address
is in the valid range or fail otherwise. The table in the figure shows for a
given instruction $i$ the result of executing it in configuration $\Phi$. The
$\fail$ and $\halt$ instructions simply fail and halt. The $\move{}{}$
instruction simply modifies the register file as requested and updates the
$\pcreg$ to the next instruction using the meta-function $\stdUpdatePc{}$.

@@

\textbf{Stack management discipline}
The precautions the needed to safely interact with an adversary have now been introduced and motivated. The stack discipline comprises the precautions. Summarized, they were:
\begin{description}[font=\normalfont\itshape]
\item[At program startup] a local $\rwlx$-capability (the stack pointer) resides in register $r_\stk$. The system does not hand out global write-local capabilities.
\item[Before returning to the adversary] clear the registers (with words we do not want to pass to the adversary) and erase any part of the stack we have used (any part of the stack that has been passed to an adversary must also be erased).
\item[Before invoking the adversary] push an activation record to the stack, create a local $\entry$-capability for the restore instructions of the activation record (the return pointer) and leave it in a register, leave a copy of the stack capability restricted to the unused part in $r_\stk$, erase the part of the stack we release control over, and clear the non-argument registers.
\item[Before invoking an adversary callback] make sure it is a global capability.
\item[When called by an adversary] make sure the alleged stack pointer has permission $\rwlx$.
\end{description}
\textbf{Reusable macro instructions} 
We define a number of reusable macros. Some of them capture the stack
management discipline and others make the examples in
Section~\ref{sec:examples} readable. The macros are:
\texttt{fetch $r$ $\var{name}$} fetches the capability related to
$\var{name}$ from the linking table and stores it in register
$r$. \texttt{push $r$}/\texttt{pop $r$} works as one would expect. All
macros that uses the stack assumes a stack pointer in register
$r_\stk$.  \texttt{prepstk $r$} is used when a callback is invoked by
the adversary and prepares the received stack pointer by checking that
it has permission $\rwlx$ and making it conform to our stack
discipline. \texttt{scall
  $r$($\overline{r_{\var{args}}}$,$\overline{r_{\var{priv}}}$)} Jumps
to the capability in register $r$ in the manner described above. That
is, it pushes local state (the contents of registers
$\overline{r_{\var{priv}}}$), stack pointer and return pointer to the
stack, sets up the necessary on stack return code and creates a return
pointer, restricts the stack pointer, clears the unused part of the
stack, clears the necessary registers and jumps to $r$. Upon return,
the private state is restored. \texttt{mclear $r$} clears all the
memory the capability in register $r$ has authority
over. \texttt{rclear $\var{regSet}$} clears all the registers in
$\var{regSet}$. \texttt{reqglob $r$} checks whether the word in
register $r$ is a global capability. \texttt{crtcls
  $\overline{(x_i,r_i)}$ $r$} allocates a closure pair where the
capability in register $r$ is used to point to the program and a new
environment is allocated (using malloc) where the contents of
$\overline{r_i}$ is stored. In the program referred to by $r$, an
implicit fetch happens when an instruction refers to $x_i$.

@@

\subsection{Encapsulation of local state}
The capability machine provides encapsulation of local state. The programs
\texttt{f1} in \figurename~\ref{fig:prog-f1} and \texttt{f2}
\figurename~\ref{fig:prog-f2} demonstrate this. They are very similar: both
store some local state, call an untrusted piece of code ($\adv$), and then test
whether the local state is unchanged. They differ in the way they do this.
Program \texttt{f1} assumes it has access to a stack pointer and uses the stack
to store its local state. It also uses the stack-based calling convention
\texttt{scall} to call the untrusted code. On the other hand, \texttt{f2} uses
malloc to allocate memory for its local state and uses an activation-record
based calling convention (described in the technical appendix) to run the
adversarial code.

@@

In order to formulate lemmas about \texttt{f1} and \texttt{f2}, we
need a way to observe whether the assertion fails. To this end, we
assume access to a flag (an address in memory). When the assertion
fails, it sets the flag to $1$ and halts. A slightly simplified
version of the correctness lemma for \texttt{f1} is:
\begin{lemma}
  \label{lem:correctness-f1}
  Let
\[
    \begin{array}{r@{}lr@{}l}
    c_{\var{adv}} & \;\defeq ((\entry,\glob),\dots) & c_{\var{stk}} & \;\defeq ((\rwlx,\local),\dots)\\
    c_{f1} & \;\defeq ((\rwx,\glob),\dots) & c_\link &\; \defeq ((\readonly,\glob),\dots)\\
    c_\malloc &\; \defeq ((\entry,\glob),\dots) & \reg& \; \in \Regs \\
    m &  \multicolumn{3}{@{}l}{\;\defeq \ms_{f1} \uplus \ms_\flag \uplus \ms_{\var{link}} \uplus \hs_\adv \uplus \ms_{\malloc} \uplus} \\
      & \multicolumn{3}{@{}l}{\phantom{\;\defeq \;}  \ms_{\var{stk}} \uplus \ms_{\var{frame}}} \\
    \end{array}
\]
  where each of the capabilities have an appropriate range of authority and pointer. Furthermore
  \begin{itemize}
  \item $\ms_{f1}$ contains $c_\link$, $c_\flag$ and the code of \texttt{f1}
  \item $\ms_\flag(\flag) = 0$
  \item $\ms_{\var{link}}$ contains $c_\adv$ and $c_\malloc$
  \item $\hs_\adv$ contains $c_\link$ and otherwise only instructions.
  \end{itemize}
  If $(\reg\update{\pcreg}{c_{f1}}\update{r_\stk}{c_\stk},m) \step^* (\halted,m')$,
  then $m'(\flag) = 0$
\end{lemma}

To prove Lemma~\ref{lem:correctness-f1}, it suffices to show that the start
configuration is capability safe (in the $\observations$ relation) for a world
with a permanent region that requires the assertion flag to be 0. We prove this
using an anti-reduction lemma by which it suffices to show that the
configuration is capability safe after some reduction steps. We then use a general lemma
for reasoning about \texttt{scall}, by which it suffices to show that (1) the
configuration that \texttt{scall} will jump to is capability safe and (2) that
the configuration just after \texttt{scall} is done cleaning up is capability
safe. We use the Fundamental Theorem to reason about the unknown code, but
notice that the Fundamental Theorem says nothing about enter capabilities and
that the adversary capability is an enter capability. Luckily the enter
capability becomes $\exec$ after the jump and then, the Fundamental Theorem
applies.

We have a similar lemma for \texttt{f2}:
\begin{lemma}
  \label{lem:correctness-f2}
  Making similar assumptions about capabilities and linking as in
  Lemma~\ref{lem:correctness-f1} but assuming no stack pointer
  if $(\reg\update{\pcreg}{c_{f2}},m) \step^* (\halted,m')$, then $m'(\flag) = 0$.
\end{lemma}

\subsection{Well-bracketed control-flow} 
Using the stack-based calling convention of \texttt{scall}, we get
well-bracketed control-flow. To illustrate this, we look at
two example programs \texttt{f3} in \figurename~\ref{fig:prog-f3} and
\texttt{g1} in \figurename~\ref{fig:prog-g1}.

In \texttt{f3} there are two calls to an adversary and in order for the
assertion in the middle to succeed, they need to be well-bracketed. If the
adversary was able to store the return pointer from the first call and invoke
it in the second call, then \texttt{f3} would have $2$ on top of its stack and
the assertion will fail. However, the security measures of our calling
convention explained in Section~\ref{sec:stack-and-return-pointer} prevent this
sort of attack: specifically, the return pointer is local, so it can only be
stored on the stack, but the part of the stack that is accessible to the
adversary is cleared before the second invocation. The following lemma shows
that there are also no other attacks that will break well-bracketedness of this
example, i.e.\ the assertion never fails. It is similar to the two previous
lemmas:
\begin{figure}[t]
  \centering
  \begin{lstlisting}
f3: push 1
    fetch $r_1$ $\adv$
    scall $r_1$($[]$,$[r_1]$)
    pop $r_2$
    assert $r_2$ 1
    push 2
    scall $r_1$($[]$,$[]$)
    halt
\end{lstlisting}
  \caption{Simple program that relies on control-flow integrity.}
  \label{fig:prog-f3}
\end{figure}

@@

The final example, in \figurename~\ref{fig:prog-g1}, is a faithful translation
of a tricky example known from the
literature~\citep{pitts_operational_1998,Dreyer:jfp12}. It consists of two
parts, \texttt{g1} and \texttt{f4}. \texttt{g1} is a closure generator that
generates closures with one variable $x$ set to $0$ in its environment and
\texttt{f4} as the program. \texttt{f4} expects one argument, a callback. The
program sets $x$ to $0$ and calls the callback. When it returns, $\texttt{f4}$
sets $x$ to $1$ and calls the callback a second time. When it returns from the
callback, it asserts that $x$ is $1$ and returns. This example is more
complicated than the previous ones because it involves a closure invoked by the
adversary and an adversary callback invoked by us. As explained in
Section~\ref{sec:stack-and-return-pointer}, this means we need to (1) take care
to properly check the stack pointer received by our closure from the adversary
(it must have the correct permission and fit with our stack discipline) and (2)
require that the adversary callback is global.

\begin{figure}[t]
  \centering
  \begin{lstlisting}
g1:  malloc $r_2$ 1
     store $r_2$ 0
     move $\pcreg$ $r_3$
     lea $r_3$ $\var{offset}$
     crtcls $[(x, r_2)]$ $r_3$
     rclear $\RegName \setminus \{\pcreg,r_0,r_1 \}$
     jmp $r_0$
f4:  reqglob $r_1$
     prepstk $r_\stk$
     store $x$ 0
     scall $r_1$($[]$,$[r_0,r_1,r_{\var{env}}]$)
     store $x$ 1
     scall $r_1$($[]$,$[r_0,r_{\var{env}}]$)
     load $r_1$ $x$
     assert $r_1$ 1
     mclear $r_\stk$
     rclear $\RegName \setminus \{r_0,\pcreg\}$
     jmp $r_0$
\end{lstlisting}
  \caption{A program that relies on well-bracketedness of
    \texttt{scall}s to function correctly. $\var{offset}$ is the
    offset to $f4$.}
  \label{fig:prog-g1}
\end{figure}

@@

In Lemma~\ref{lem:correctness-g1} the order of control has been inverted
compared to the previous lemmas. In this lemma, the adversary assumes control
first with a capability for the closure creator \texttt{g1}. This means we need
to check that all arguments are safe to use and that we clean up before
returning in the end. The inversion of control poses an interesting challenge
when it comes to reasoning about the adversary's local state during the
execution of \texttt{f4}. During the execution of \texttt{f4} and the callbacks,
the adversary should not be able to rely on its local state from before the call
of \texttt{f4}. This is easily done by revoking all the temporary regions of the
world given at the start of \texttt{f4}. However, when \texttt{f4} returns, the
adversary is again allowed to rely on its old local state which means that we
need to guarantee that the local state is unchanged. This is important because
the return pointer that \texttt{f4} receives from the adversary may be local,
and the adversary is allowed to allocate the activation record on the stack
(just like we do) so that they can store and recover his old stack pointer after
\texttt{f4} returns.

In Figure~\ref{fig:worlds-in-pf-g1}, we illustrate how we accomodate this in the
proof of Lemma~\ref{lem:correctness-g1} by constructing appropriate worlds for
all the situations where control is passed to the adversary. The potentially
local return pointer received by \texttt{f4} from the adversary is safe in
$W_1$, so it can only be used in public future worlds. Let us see how $W_6$ is
constructed as a public future world. The worlds $W_2$ and $W_4$ are constructed
by revoking all temporary regions and adding a $\iota^\pwl$ region for the stack
we pass away and static temporary regions (a region that accepts exactly one
memory) for our local stack as well as the adversary's local stack. The given
worlds $W_3$ and $W_5$ are public future worlds in which the adversary chooses
to return to us. None of these worlds can have masked the temporary regions of
$W_1$ with permanent regions. $W_6$ is constructed by revoking the temporary
regions of $W_5$ and reinstating the temporary regions of $W_1$. This makes
$W_6$ a public future world of $W_1$. It is therefore safe to use the return
pointer to return to the adversary, since we have restored validity of any local
state they might have stashed away.

@@

\textbf{Formulating control flow integrity} The attentive reader may notice
that, while we claim that our stack and return pointer discipline enforces
control flow integrity, we do not prove a general theorem to show that this is
the case. The reason for this is that it is not clear what such a theorem should
look like. Formulations in terms of a control flow graph, like the one by
\cite{abadi_control-flow_2005}, do not take into account temporal properties,
like the well-bracketedness that Example~\texttt{g1} relies on.\footnote{It is a
  bit unfortunate that the general term ``control flow integrity'' is used for a
  property that only captures partial correctness of control flow.} In fact, our
examples show that our logical relation imply a form of control flow integrity
that is stronger than such formulations, although this is not made very
explicit. As future work, we are considering to look at a more explicit and
useful way to formalise well-bracketed control flow. The idea would be to define
a variant of our capability machine with built-in call and return instructions
and well-bracketed control flow built-in to the operational semantics, and then
prove that compiling such code to our machine using our stack and return pointer
discipline is fully abstract~\cite{abadi_protection_1998}, which roughly means
that the lower-level machine does not allow interactions with the compiled code
that do not correspond to an interaction with the original code in the
high-level machine.

>>>>

\usepackage{cite}
\usepackage{amsmath,amsfonts,amssymb,amsthm}
\usepackage{thmtools}
\usepackage{stmaryrd}
\usepackage{natbib}
\usepackage{url}
\usepackage{array}
\usepackage{arydshln}
\usepackage{ifthen}
\usepackage{ifpdf}
\usepackage{verbatim}
\usepackage{mathpartir}
\usepackage{listings}
\usepackage{hyperref}
\usepackage{multicol}
\usepackage{natbib}
\lstset{
  basicstyle=\footnotesize\ttfamily,
  columns=fullflexible,
  keepspaces=true,
  mathescape
}

@@

\begin{comment}
  \begin{figure}
    \centering
    \begin{mathpar}
      \Phi \rightarrow
      \begin{cases}
        \sem{\decode(n)}(\Phi) & \arraycolsep=0pt
        \begin{array}[t]{l}
          \text{\footnotesize{if $\memreg(\pcreg) = \stdcap$}}\\
          \;\;\text{\footnotesize{and $\start \leq \addr \leq \addrend$ and $\memheap(\addr) = n$}}\\
          \;\;\text{\footnotesize{and $\perm \in \{ \exec,\rwx, \rwlx \}$} }
        \end{array}\\
        \failed                & \text{\footnotesize{otherwise}}
      \end{cases}
      \and \updatePcPerm{w} =
      \begin{cases}
        ((\exec,\gl),\start,\addrend,\addr) & \text{\footnotesize{if $w = ((\entry,\gl),\start,\addrend,\addr)$}}\\
        w & \text{\footnotesize{otherwise}}
      \end{cases}\\
      \and \stdUpdatePc{\Phi} =
      \begin{cases}
        \updateReg{\pcreg}{\var{newPc}} & \arraycolsep=0pt
        \begin{array}[t]{l}
          \text{\footnotesize{if $\memreg(\pcreg) = \stdcap$}}\\ 
          \;\;\text{\footnotesize{and $\var{newPc} = ((\perm,\gl),\start,\addrend,\addr + 1)$}}
        \end{array}\\
        \failed & \text{\footnotesize{otherwise}}
      \end{cases}
      \and \inferrule{ } { \sem{\fail}(\Phi) = \failed } \and
      \inferrule{ } { \sem{\halt}(\Phi) = (\halted,\memheap) } \and
      \inferrule{ } { \sem{\move{\refreg{r_1}}{r_2}}(\Phi) =
        \stdUpdatePc{}(\updateReg{r_1}{\memreg(r_2)}) } \and
      \inferrule{ \memreg(r_2) = \stdcap \\
        \perm \in \{ \rwx, \rwlx, \exec, \readwrite,
        \readwritel, \readonly \}\\
        \start \leq \addr \leq \addrend } {
        \sem{\load{\refreg{r_1}}{\refheap{r_2}}}(\Phi) =
        \updateReg{r_1}{\memheap(\addr)}} \and
      \inferrule{ \memreg(r_2) = \stdcap[\permp]\\
        \permp' = \decodePermPair{\memreg(r_2)}\\
        \permp'\sqsubseteq \permp } {
        \sem{\restricttwo{\refreg{r_1}}{r_2}}(\Phi) = \\
        \stdUpdatePc{}(\updateReg{r_1}{(\permp',\start,\addrend,\addr)})
      } \and \inferrule{ \memreg(r_2) = ((\_,\_),\_,\_,\addr) } {
        \sem{\geta{\refreg{r_1}}{\refreg{r_2}}}(\Phi) =
        \stdUpdatePc{\updateReg{r_1}{\addr}}} \and \inferrule{
        \var{newPc} = \updatePcPerm{\memreg(\lv)} } {
        \sem{\jmp{\lv}}(\Phi) = \updateReg{\pcreg}{\var{newPc}} } \and
      \inferrule{ \memreg(r_1) = \stdcap\\
        \start \leq \addr \leq \addrend\\
        \perm \in \{ \rwx, \rwlx, \readwrite, \readwritel\}\\
        \var{w} = \memreg(r_2) \\
        \var{w} = ((\_,\local),\_,\_,\_)\Rightarrow\perm \in
        \{\rwlx,\readwritel \} } {
        \sem{\store{\refheap{r_1}}{\refreg{r_2}}}(\Phi) =
        \stdUpdatePc{\updateHeap{\addr}{\var{w}}} } \and
    \end{mathpar}
    \caption{An excerpt from the operational semantics.}
    \label{fig:op-sem}
  \end{figure}
\end{comment}
\figurename~\ref{fig:op-sem} shows an excerpt of the operational semantics containing a number of representative
instructions. Essentially, a configuration $\Phi$ will either decode and execute
the instruction at $\memreg{\pcreg}$ if it is executable and the current address
is in the valid range or fail otherwise. The table in the figure shows for a
given instruction $i$ the result of executing it in configuration $\Phi$. The
$\fail$ and $\halt$ instructions simply fail and halt. The $\move{}{}$
instruction simply modifies the register file as requested and updates the
$\pcreg$ to the next instruction using the meta-function $\stdUpdatePc{}$.

@@

\textbf{Stack management discipline}
The precautions the needed to safely interact with an adversary have now been introduced and motivated. The stack discipline comprises the precautions. Summarized, they were:
\begin{description}[font=\normalfont\itshape]
\item[At program startup] a local $\rwlx$-capability (the stack pointer) resides in register $r_\stk$. The system does not hand out global write-local capabilities.
\item[Before returning to the adversary] clear the registers (with words we do not want to pass to the adversary) and erase any part of the stack we have used (any part of the stack that has been passed to an adversary must also be erased).
\item[Before invoking the adversary] push an activation record to the stack, create a local $\entry$-capability for the restore instructions of the activation record (the return pointer) and leave it in a register, leave a copy of the stack capability restricted to the unused part in $r_\stk$, erase the part of the stack we release control over, and clear the non-argument registers.
\item[Before invoking an adversary callback] make sure it is a global capability.
\item[When called by an adversary] make sure the alleged stack pointer has permission $\rwlx$.
\end{description}
\textbf{Reusable macro instructions} 
We define a number of reusable macros. Some of them capture the stack
management discipline and others make the examples in
Section~\ref{sec:examples} readable. The macros are:
\texttt{\footnotesize{fetch $r$ $\var{name}$}} fetches the capability related to
$\var{name}$ from the linking table and stores it in register
$r$. \texttt{\footnotesize{push $r$}}/\texttt{\footnotesize{pop $r$}} works as one would expect. All
macros that uses the stack assumes a stack pointer in register
$r_\stk$.  \texttt{\footnotesize{prepstk $r$}} is used when a callback is invoked by
the adversary and prepares the received stack pointer by checking that
it has permission $\rwlx$ and making it conform to our stack
discipline. \texttt{\footnotesize{scall $r$($\overline{r_{\var{args}}}$,$\overline{r_{\var{priv}}}$)}} Jumps
to the capability in register $r$ in the manner described above. That
is, it pushes local state (the contents of registers
$\overline{r_{\var{priv}}}$), stack pointer and return pointer to the
stack, sets up the necessary on stack return code and creates a return
pointer, restricts the stack pointer, clears the unused part of the
stack, clears the necessary registers and jumps to $r$. Upon return,
the private state is restored. \texttt{\footnotesize{mclear $r$}} clears all the
memory the capability in register $r$ has authority
over. \texttt{\footnotesize{rclear $\var{regSet}$}} clears all the registers in
$\var{regSet}$. \texttt{\footnotesize{reqglob $r$}} checks whether the word in
register $r$ is a global capability. \texttt{\footnotesize{crtcls
  $\overline{(x_i,r_i)}$ $r$}} allocates a closure pair where the
capability in register $r$ is used to point to the program and a new
environment is allocated (using malloc) where the contents of
$\overline{r_i}$ is stored. In the program referred to by $r$, an
implicit fetch happens when an instruction refers to $x_i$.

@@

\subsection{Encapsulation of local state}
The capability machine provides encapsulation of local state. The programs
\texttt{\footnotesize{f1}} in \figurename~\ref{fig:prog-f1} and \texttt{\footnotesize{f2}}
\figurename~\ref{fig:prog-f2} demonstrate this. They are very similar: both
store some local state, call an untrusted piece of code ($\adv$), and then test
whether the local state is unchanged. They differ in the way they do this.
Program \texttt{\footnotesize{f1}} assumes it has access to a stack pointer and uses the stack
to store its local state. It also uses the stack-based calling convention
\texttt{\footnotesize{scall}} to call the untrusted code. On the other hand, \texttt{\footnotesize{f2}} uses
malloc to allocate memory for its local state and uses an activation-record
based calling convention (described in the technical appendix) to run the
adversarial code.

@@

In order to formulate lemmas about \texttt{\footnotesize{f1}} and \texttt{\footnotesize{f2}}, we
need a way to observe whether the assertion fails. To this end, we
assume access to a flag (an address in memory). When the assertion
fails, it sets the flag to $1$ and halts. A slightly simplified
version of the correctness lemma for \texttt{\footnotesize{f1}} is:
\begin{lemma}
  \label{lem:correctness-f1}
  Let
\[
    \begin{array}{r@{}lr@{}l}
    c_{\var{adv}} & \;\defeq ((\entry,\glob),\dots) & c_{\var{stk}} & \;\defeq ((\rwlx,\local),\dots)\\
    c_{f1} & \;\defeq ((\rwx,\glob),\dots) & c_\link &\; \defeq ((\readonly,\glob),\dots)\\
    c_\malloc &\; \defeq ((\entry,\glob),\dots) & \reg& \; \in \Regs \\
    m &  \multicolumn{3}{@{}l}{\;\defeq \ms_{f1} \uplus \ms_\flag \uplus \ms_{\var{link}} \uplus \hs_\adv \uplus \ms_{\malloc} \uplus} \\
      & \multicolumn{3}{@{}l}{\phantom{\;\defeq \;}  \ms_{\var{stk}} \uplus \ms_{\var{frame}}} \\
    \end{array}
\]
  where each of the capabilities have an appropriate range of authority and pointer. Furthermore
  \begin{itemize}
  \item $\ms_{f1}$ contains $c_\link$, $c_\flag$ and the code of \texttt{\footnotesize{f1}}
  \item $\ms_\flag(\flag) = 0$
  \item $\ms_{\var{link}}$ contains $c_\adv$ and $c_\malloc$
  \item $\hs_\adv$ contains $c_\link$ and otherwise only instructions.
  \end{itemize}
  If $(\reg\update{\pcreg}{c_{f1}}\update{r_\stk}{c_\stk},m) \step^* (\halted,m')$,
  then $m'(\flag) = 0$
\end{lemma}

To prove Lemma~\ref{lem:correctness-f1}, it suffices to show that the start
configuration is capability safe (in the $\observations$ relation) for a world
with a permanent region that requires the assertion flag to be 0. We prove this
using an anti-reduction lemma by which it suffices to show that the
configuration is capability safe after some reduction steps. We then use a general lemma
for reasoning about \texttt{\footnotesize{scall}}, by which it suffices to show that (1) the
configuration that \texttt{\footnotesize{scall}} will jump to is capability safe and (2) that
the configuration just after \texttt{\footnotesize{scall}} is done cleaning up is capability
safe. We use the Fundamental Theorem to reason about the unknown code, but
notice that the Fundamental Theorem says nothing about enter capabilities and
that the adversary capability is an enter capability. Luckily the enter
capability becomes $\exec$ after the jump and then, the Fundamental Theorem
applies.

We have a similar lemma for \texttt{\footnotesize{f2}}:
\begin{lemma}
  \label{lem:correctness-f2}
  Making similar assumptions about capabilities and linking as in
  Lemma~\ref{lem:correctness-f1} but assuming no stack pointer
  if $(\reg\update{\pcreg}{c_{f2}},m) \step^* (\halted,m')$, then $m'(\flag) = 0$.
\end{lemma}

\subsection{Well-bracketed control-flow} 
Using the stack-based calling convention of \texttt{\footnotesize{scall}}, we get
well-bracketed control-flow. To illustrate this, we look at
two example programs \texttt{\footnotesize{f3}} in \figurename~\ref{fig:prog-f3} and
\texttt{\footnotesize{g1}} in \figurename~\ref{fig:prog-g1}.

In \texttt{\footnotesize{f3}} there are two calls to an adversary and in order for the
assertion in the middle to succeed, they need to be well-bracketed. If the
adversary was able to store the return pointer from the first call and invoke
it in the second call, then \texttt{\footnotesize{f3}} would have $2$ on top of its stack and
the assertion will fail. However, the security measures of our calling
convention explained in Section~\ref{sec:stack-and-return-pointer} prevent this
sort of attack: specifically, the return pointer is local, so it can only be
stored on the stack, but the part of the stack that is accessible to the
adversary is cleared before the second invocation. The following lemma shows
that there are also no other attacks that will break well-bracketedness of this
example, i.e.\ the assertion never fails. It is similar to the two previous
lemmas:
\begin{figure}[t]
  \centering
  \begin{lstlisting}
f3: push 1
    fetch $r_1$ $\adv$
    scall $r_1$($[]$,$[r_1]$)
    pop $r_2$
    assert $r_2$ 1
    push 2
    scall $r_1$($[]$,$[]$)
    halt
\end{lstlisting}
  \caption{Simple program that relies on control-flow integrity.}
  \label{fig:prog-f3}
\end{figure}

@@

The final example, in \figurename~\ref{fig:prog-g1}, is a faithful translation
of a tricky example known from the
literature~\citep{pitts_operational_1998,Dreyer:jfp12}. It consists of two
parts, \texttt{\footnotesize{g1}} and \texttt{\footnotesize{f4}}. \texttt{\footnotesize{g1}} is a closure generator that
generates closures with one variable $x$ set to $0$ in its environment and
\texttt{\footnotesize{f4}} as the program. \texttt{\footnotesize{f4}} expects one argument, a callback. The
program sets $x$ to $0$ and calls the callback. When it returns, \texttt{\footnotesize{f4}}
sets $x$ to $1$ and calls the callback a second time. When it returns from the
callback, it asserts that $x$ is $1$ and returns. This example is more
complicated than the previous ones because it involves a closure invoked by the
adversary and an adversary callback invoked by us. As explained in
Section~\ref{sec:stack-and-return-pointer}, this means we need to (1) take care
to properly check the stack pointer received by our closure from the adversary
(it must have the correct permission and fit with our stack discipline) and (2)
require that the adversary callback is global.

\begin{figure}[t]
  \centering
  \begin{lstlisting}
g1:  malloc $r_2$ 1
     store $r_2$ 0
     move $\pcreg$ $r_3$
     lea $r_3$ $\var{offset}$
     crtcls $[(x, r_2)]$ $r_3$
     rclear $\RegName \setminus \{\pcreg,r_0,r_1 \}$
     jmp $r_0$
f4:  reqglob $r_1$
     prepstk $r_\stk$
     store $x$ 0
     scall $r_1$($[]$,$[r_0,r_1,r_{\var{env}}]$)
     store $x$ 1
     scall $r_1$($[]$,$[r_0,r_{\var{env}}]$)
     load $r_1$ $x$
     assert $r_1$ 1
     mclear $r_\stk$
     rclear $\RegName \setminus \{r_0,\pcreg\}$
     jmp $r_0$
\end{lstlisting}
  \caption{A program that relies on well-bracketedness of
    \texttt{scall}s to function correctly. $\var{offset}$ is the
    offset to \texttt{f4}.}
  \label{fig:prog-g1}
\end{figure}

@@

In Lemma~\ref{lem:correctness-g1} the order of control has been inverted
compared to the previous lemmas. In this lemma, the adversary assumes control
first with a capability for the closure creator \texttt{\footnotesize{g1}}. This means we need
to check that all arguments are safe to use and that we clean up before
returning in the end. The inversion of control poses an interesting challenge
when it comes to reasoning about the adversary's local state during the
execution of \texttt{\footnotesize{f4}}. During the execution of \texttt{\footnotesize{f4}} and the callbacks,
the adversary should not be able to rely on its local state from before the call
of \texttt{\footnotesize{f4}}. This is easily done by revoking all the temporary regions of the
world given at the start of \texttt{\footnotesize{f4}}. However, when \texttt{\footnotesize{f4}} returns, the
adversary is again allowed to rely on its old local state which means that we
need to guarantee that the local state is unchanged. This is important because
the return pointer that \texttt{\footnotesize{f4}} receives from the adversary may be local,
and the adversary is allowed to allocate the activation record on the stack
(just like we do) so that they can store and recover his old stack pointer after
\texttt{\footnotesize{f4}} returns.

In Figure~\ref{fig:worlds-in-pf-g1}, we illustrate how we accomodate this in the
proof of Lemma~\ref{lem:correctness-g1} by constructing appropriate worlds for
all the situations where control is passed to the adversary. The potentially
local return pointer received by \texttt{\footnotesize{f4}} from the adversary is safe in
$W_1$, so it can only be used in public future worlds. Let us see how $W_6$ is
constructed as a public future world. The worlds $W_2$ and $W_4$ are constructed
by revoking all temporary regions and adding a $\iota^\pwl$ region for the stack
we pass away and static temporary regions (a region that accepts exactly one
memory) for our local stack as well as the adversary's local stack. The given
worlds $W_3$ and $W_5$ are public future worlds in which the adversary chooses
to return to us. None of these worlds can have masked the temporary regions of
$W_1$ with permanent regions. $W_6$ is constructed by revoking the temporary
regions of $W_5$ and reinstating the temporary regions of $W_1$. This makes
$W_6$ a public future world of $W_1$. It is therefore safe to use the return
pointer to return to the adversary, since we have restored validity of any local
state they might have stashed away.

@@

\textbf{Formulating control flow integrity} The attentive reader may notice
that, while we claim that our stack and return pointer discipline enforces
control flow integrity, we do not prove a general theorem to show that this is
the case. The reason for this is that it is not clear what such a theorem should
look like. Formulations in terms of a control flow graph, like the one by
\cite{abadi_control-flow_2005}, do not take into account temporal properties,
like the well-bracketedness that Example~\texttt{\footnotesize{g1}} relies on.\footnote{It is a
  bit unfortunate that the general term ``control flow integrity'' is used for a
  property that only captures partial correctness of control flow.} In fact, our
examples show that our logical relation imply a form of control flow integrity
that is stronger than such formulations, although this is not made very
explicit. As future work, we are considering to look at a more explicit and
useful way to formalise well-bracketed control flow. The idea would be to define
a variant of our capability machine with built-in call and return instructions
and well-bracketed control flow built-in to the operational semantics, and then
prove that compiling such code to our machine using our stack and return pointer
discipline is fully abstract~\cite{abadi_protection_1998}, which roughly means
that the lower-level machine does not allow interactions with the compiled code
that do not correspond to an interaction with the original code in the
high-level machine.
